CF Registry Analysis

### 2022-12-26 Christine Chan/CF Registry/cf_registry_analysis.qmd
both CFRD definitions done (I think)

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    page-layout: full
editor: visual
---

@@

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
# Add age
ages <- annual %>% arrange(eDWID,ReviewYear) %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  group_by(eDWID) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])))
encounter = left_join(encounter,ages,by = "eDWID")
encounter$`Age Group` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd <- encounter %>%
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  slice_min(reviewyear) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd = reviewyear)
encounter <- left_join(encounter, first_cfrd, by = c("eDWID", "Age in 2018"))
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(cfrd_ogtt = case_when(
    reviewyear < first_cfrd ~ "No",
    reviewyear >= first_cfrd ~ "Yes"
  )) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
```

```{r}
#| label: tbl-incidence
#| tbl-cap: "Year-on-Year Incidence of CFRD"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,is.na(first_cfrd) | first_cfrd == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>%
    select(eDWID,reviewyear,encounternum,`Age Group`,
           twohour_bloodglucose,cfrd_ogtt,first_cfrd)
  inc = inc %>% group_by(cfrd_ogtt) %>% count(`Age Group`,.drop = F)
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_ogtt=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

  new_cases_10 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "< 10"])
  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
  inc_10 = new_cases_10/(at_risk_10/py)

  new_cases_10_18 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "10 - 18"])
  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)

  new_cases_19_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "19 - 30"])
  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)

  new_cases_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "> 30"])
  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
  inc_30 = new_cases_30/(at_risk_30/py)

@@

In @tbl-incidence, incidence per year is calculated as the number of new CFRD diagnoses per 1,000 person-years of those at risk (excluding previously identified cases).

```{r}
#| label: tbl-prevalence
#| tbl-cap: "Yearly Prevalence of CFRD"
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(highest_2_hour_bg = suppressWarnings(max(twohour_bloodglucose,na.rm = T)),
            cfrd = max(cfrd_ogtt),
            cfrd_status = suppressWarnings(max(cfrd_status,na.rm = T)),
            .groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(`Prevalence by OGTT` = mean(as.numeric(cfrd)-1)*100,
            `Prevalence by CFFPR` = mean(cfrd_status == 3,na.rm = T)*100) %>%
  kable(.,digits = 3)
```

In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.

>>>>

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    code-fold: true
    self-contained: true
    page-layout: full
editor: source
---

@@

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
# Add age
ages <- annual %>% arrange(eDWID,ReviewYear) %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  group_by(eDWID) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])))
encounter = left_join(encounter,ages,by = "eDWID")
encounter$`Age Group` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
```

```{r}
#| label: tbl-incidence_cffpr
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>%  
    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>% 
    select(eDWID,`Age Group`,cfrd_cffpr)
  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "< 10"])
  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
  inc_10 = new_cases_10/(at_risk_10/py)

  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "10 - 18"])
  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)

  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "19 - 30"])
  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)

  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "> 30"])
  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
  inc_30 = new_cases_30/(at_risk_30/py)

@@

In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.

```{r}
#| label: tbl-prevalence_cffpr
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
  kable(.,digits = 3)
```

In @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-incidence_both
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR and OGTT"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>%
    select(eDWID,reviewyear,encounternum,`Age Group`,
           twohour_bloodglucose,cfrd_both,first_cfrd_both)
  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_both=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

  new_cases_10 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "< 10"])
  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
  inc_10 = new_cases_10/(at_risk_10/py)

  new_cases_10_18 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "10 - 18"])
  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)

  new_cases_19_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "19 - 30"])
  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)

  new_cases_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "> 30"])
  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
  inc_30 = new_cases_30/(at_risk_30/py)

In @tbl-incidence_both, incidence per year is calculated as the number of new CFRD diagnoses (by either registry definition or OGTT, whichever happened first) per 1,000 person-years of those at risk, excluding previously identified cases.

```{r}
#| label: tbl-prevalence_both
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
  kable(.,digits = 3)
```

In @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-screening
#| tbl-cap: "Yearly Screening Rates"

```

In @tbl-screening, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded. Screening rate is calculated as the number of people with OGTTs divided by the total number of people seen in clinic that year.
